---
title: "statistical_computing_project4"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(DT)
library(jsonlite)
library(httr)
library(here)
library(purrr)
library(tidyverse)
library(tidycensus)
census_api_key("dbd69da6dd6f0e9d267b0cbdd2168231ea942714",overwrite=TRUE, install = TRUE)
readRenviron("~/.Renviron")

racevars <- c(White = "P2_005N", 
              Black = "P2_006N", 
              Asian = "P2_008N", 
              Hispanic = "P2_002N")

if(!file.exists(here("data","bal_population.RDS"))){
  baltimore <- get_decennial(
    geography = "tract",
    variables = racevars,
    state = '24',
    county = '005',
    geometry = TRUE,
    summary_var = "P2_001N",
    year = 2020
  )
  walk(.x = baltimore,
                  .f = ~saveRDS(.x, 
                    file = "C:/Users/Lenovo/Documents/project1/data/bal_population.RDS"))
}

if(!file.exists(here("data","alameda_population.RDS"))){
  alameda <- get_decennial(
    geography = "tract",
    variables = racevars,
    state = '06',
    county = '001',
    geometry = TRUE,
    summary_var = "P2_001N",
    year = 2020
  ) 
  walk(.x = alameda,
                  .f = ~saveRDS(.x, 
                    file = "C:/Users/Lenovo/Documents/project1/data/alameda_population.RDS"))
}

baltimore <- readRDS(here("data","bal_population.RDS"))
alameda <- readRDS(here("data","alameda_population.RDS"))
```

# About

**Motivation and Overview**

This dashboard manages to explore the difference of population in two counties that I lived in.


**A video explaining this dashboard**

# Data

The population data of baltimore I used is obtained by [tidycensus API](https://walker-data.com/tidycensus/) and shown on the left. The data of berkeley city is similar on the right.

***

Column {.tabset}
-------------------------------------


### Baltimore data

```{r echo=FALSE}
renderDataTable({
  datatable(baltimore,
                caption = tags$caption(style = 'caption-side: Left; text-align: Left;'),
                options = list(autoWidth = TRUE,
                               pageLength = 10,
                               scroller = TRUE,
                               scrollY = '750px'))
})
```


### Alameda data

```{r echo=FALSE}
renderDataTable({
  datatable(alameda,
                caption = tags$caption(
                  style = 'caption-side: top; text-align: Left;'),
                options = list(autoWidth = TRUE,
                               pageLength = 10,
                               scroller = TRUE,
                               scrollY = '450px'))
})
```

# Total population distribution


Column {data-width=500}
-------------------------------------

### Baltimore
```{r echo=FALSE}
baltimore %>%
  ggplot(aes(fill = summary_value)) +
  geom_sf(color = NA) +
  theme_void() + 
  scale_fill_viridis_c() + 
  labs(fill = "% of population\n(2020 Census)",
       title = "Population map in baltimore county")
```


Column {data-width=500}
-------------------------------------
### Alameda

```{r echo=FALSE}
alameda %>%
  ggplot(aes(fill = summary_value)) +
  geom_sf(color = NA) +
  theme_void() + 
  scale_fill_viridis_c() + 
  labs(fill = "% of population\n(2020 Census)",
       title = "Population map in alameda county")
```

# Population distribution by race

Column {.tabset}
-------------------------------------

### Baltimore

```{r echo=FALSE}
baltimore %>%
  mutate(percent = 100 * (value / summary_value)) %>%
  ggplot(aes(fill = percent)) +
  facet_wrap(~variable) +
  geom_sf(color = NA) +
  theme_void() + 
  scale_fill_viridis_c() + 
  labs(fill = "% of population\n(2020 Census)",
       title = "Population map by races in baltimore county")

```


### Alameda

```{r echo=FALSE}
alameda %>%
  mutate(percent = 100 * (value / summary_value)) %>%
  ggplot(aes(fill = percent)) +
  facet_wrap(~variable) +
  geom_sf(color = NA) +
  theme_void() + 
  scale_fill_viridis_c() + 
  labs(fill = "% of population\n(2020 Census)",
       title = "Population map by races in alameda county")
```

# Difference of population by race


Column {.sidebar}
-------------------------------------

The races that you are interested in.

```{r echo=FALSE}
checkboxGroupInput("counties", label = "Choose counties",
            choices = c("Asian" = "Asian",
                        "Black" = "Black",
                        "Hispanic" = "Hispanic",
                        "White" = "White"),
            selected = c("Asian", "White"))
```

Row {.tabset}
-----------------------------------------------------------------------

### Difference of population by race between and Baltimore and Alameda (Violin plot)
```{r echo=FALSE}
baltimore <- baltimore %>%
  mutate(county = "baltimore")

alameda_part1 <- alameda

renderPlot({
  alameda_part1 %>%
    mutate(county = "alameda") %>%
    bind_rows(baltimore) %>%
    filter(variable %in% input$counties) %>%
    ggplot(aes(x=county, y=value, fill=variable)) +
    geom_violin() +
    ylim(0,4500) +
    labs(x = "city", y = "population", fill = "race",
       title = "Violin plot of population by race")
})
```

### Difference of population by race between and Baltimore and Alameda (Boxplot plot)
```{r echo=FALSE}
baltimore <- baltimore %>%
  mutate(county = "baltimore")

alameda_part2 <- alameda

renderPlot({
  alameda_part2 %>%
    mutate(county = "alameda") %>%
    bind_rows(baltimore) %>%
    filter(variable %in% input$counties) %>%
    ggplot(aes(x=county, y=value, fill=variable)) +
    geom_boxplot() +
    ylim(0,4500) +
    labs(x = "city", y = "population", fill = "race",
       title = "Boxplot of population by race")
})
```


# Difference of population density by race


Column {.sidebar}
-------------------------------------

The races that you are interested in.

```{r echo=FALSE}
checkboxGroupInput("mycounties", label = "Choose counties",
            choices = c("Asian" = "Asian",
                        "Black" = "Black",
                        "Hispanic" = "Hispanic",
                        "White" = "White"),
            selected = c("Asian", "White"))
```

Row {.tabset}
-----------------------------------------------------------------------

### Difference of population by race between and Baltimore and Alameda (Violin plot)
```{r echo=FALSE}
baltimore <- baltimore %>%
  mutate(county = "baltimore")

alameda_part1 <- alameda

renderPlot({
  alameda_part1 %>%
    mutate(county = "alameda") %>%
    bind_rows(baltimore) %>%
    filter(variable %in% input$mycounties) %>%
    mutate(percent = 100 * (value / summary_value)) %>%
    ggplot(aes(x=county, y = percent, fill=variable)) +
    geom_violin() +
    labs(x = "city", y = "population density", fill = "race",
       title = "Violin plot of population density by race")
})
```

### Difference of population by race between and Baltimore and Alameda (Boxplot plot)
```{r echo=FALSE}
baltimore <- baltimore %>%
  mutate(county = "baltimore")

alameda_part2 <- alameda

renderPlot({
  alameda_part2 %>%
    mutate(county = "alameda") %>%
    bind_rows(baltimore) %>%
    filter(variable %in% input$mycounties) %>%
    mutate(percent = 100 * (value / summary_value)) %>%
    ggplot(aes(x=county, y=percent, fill=variable)) +
    geom_boxplot() +
    labs(x = "city", y = "population density", fill = "race",
       title = "Boxplot of population density by race")
})
```

# Analysis in project3

Row
-------------------------------------
    
### Population map
    
```{r echo=FALSE}
baltimore %>%
  mutate(percent = 100 * (value / summary_value)) %>%
  ggplot(aes(fill = percent)) +
  facet_wrap(~variable) +
  geom_sf(color = NA) +
  theme_void() + 
  scale_fill_viridis_c() + 
  labs(fill = "% of population\n(2020 Census)",
       title = "Population map of the number of races in baltimore county", 
       subtitle = "Baltimore has a relatively small Asian population",
       caption = "Credit to: Yangzhenyu Gao") +
  theme(plot.title = element_text(color = "dark blue", hjust = 0.5),
        plot.caption = element_text(color ="blue"),
        plot.subtitle = element_text(color = "blue"))

alameda %>%
  mutate(percent = 100 * (value / summary_value)) %>%
  ggplot(aes(fill = percent)) +
  facet_wrap(~variable) +
  geom_sf(color = NA) +
  theme_void() + 
  scale_fill_viridis_c() + 
  labs(fill = "% of population\n(2020 Census)",
       title = "Population map of the number of races in alameda county", 
       subtitle = "Baltimore has a relatively small Black population",
       caption = "Credit to: Yangzhenyu Gao") +
  theme(plot.title = element_text(color = "dark blue", hjust = 0.5),
        plot.caption = element_text(color ="blue"),
        plot.subtitle = element_text(color = "blue"))
```


Row
-------------------------------------
    
### Difference of population by race 
    
```{r echo=FALSE}
baltimore <- baltimore %>%
  mutate(county = "baltimore")

alameda_part3 <- alameda

alameda_part3 %>%
  mutate(county = "alameda") %>%
  bind_rows(baltimore) %>%
  ggplot(aes(x=county, y=value, fill=variable)) +
  geom_violin() +
  ylim(0,4500) +
  labs(x = "city", y = "population", fill = "race",
       title = "Violin plot of population", 
       subtitle = "The population diversity of Baltimore is relatively small",
       caption = "Credit to: Yangzhenyu Gao") +
  theme(plot.title = element_text(color = "dark blue", hjust = 0.5),
        plot.caption = element_text(color ="blue"),
        plot.subtitle = element_text(color = "blue"))

```
    
### Difference of population density by race

```{r echo=FALSE}
baltimore <- baltimore %>%
  mutate(percent = 100 * (value / summary_value))

alameda_part4 <- alameda

alameda_part4 %>%
  mutate(county = "alameda") %>%
  mutate(percent = 100 * (value / summary_value)) %>%
  bind_rows(baltimore) %>%
  drop_na() %>%
  select(county,variable,percent) %>%
  rename(density = percent) %>%
  ggplot(aes(x=county, y=density, fill=variable)) +
  geom_boxplot() +
  labs(x = "city", y = "population density", fill = "race",
       title = "Boxplot of population density", 
       subtitle = "The population diversity of alameda is relatively big",
       caption = "Credit to: Yangzhenyu Gao") +
  theme(plot.title = element_text(color = "dark blue", hjust = 0.5),
        plot.caption = element_text(color ="blue"),
        plot.subtitle = element_text(color = "blue"))

```


